% Copyright (c) 2023 Carl Martin Ludvig Sinander.

% This program is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.

% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
% GNU General Public License for more details.

% You should have received a copy of the GNU General Public License
% along with this program. If not, see <https://www.gnu.org/licenses/>.

%                                   _     _
%    _ __  _ __ ___  __ _ _ __ ___ | |__ | | ___
%   | '_ \| '__/ _ \/ _` | '_ ` _ \| '_ \| |/ _ \
%   | |_) | | |  __/ (_| | | | | | | |_) | |  __/
%   | .__/|_|  \___|\__,_|_| |_| |_|_.__/|_|\___|
%   |_|


%%% bug catcher
\RequirePackage[l2tabu,orthodox]{nag}

%%% document class
\documentclass[11pt,letterpaper,reqno,oneside]{book}

%%% settings
\input{preamble.tex}

%%% for \widthof
\usepackage{calc}

% break DOIs in bibliography
\setcounter{biburlnumpenalty}{100}

% raise hypertargets above baseline
\makeatletter
	\newcommand{\hyperdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
\makeatother

% % watermark on every page (for version control)
% \usepackage[anchor=ll,pos={0.1cm,0.5cm},fontsize=0.2cm,angle=0,alignment=l]{draftwatermark}
% \SetWatermarkText{\normalfont\hspace{0.37em}version:\\\normalfont {\datestyle\today}\\\normalfont\hspace{0.37em}at {\currenttime}}


%%% bibliography
\addbibresource{bibl.bib}


%______________________________________________________________________________




%    _____ _ _   _
%   |_   _(_) |_| | ___
%     | | | | __| |/ _ \
%     | | | | |_| |  __/
%     |_| |_|\__|_|\___|


\title{\scshape Topics in mechanism design}

\author{Ludvig Sinander \\
University of Oxford}

\date{\small This version: 6 November 2023}

% \date{\emph{version:} {\datestyle\today} at {\currenttime} \\ \hspace{0pt} \\ \hspace{0pt} \\ \bfseries please report typos!}

\makeatletter
	\AtBeginDocument{ \hypersetup{
		pdftitle = {Topics in mechanism design},
		pdfauthor = {Ludvig Sinander}
		} }
\makeatother



%______________________________________________________________________________




%    ____                                        _
%   |  _ \  ___   ___ _   _ _ __ ___   ___ _ __ | |_
%   | | | |/ _ \ / __| | | | '_ ` _ \ / _ \ '_ \| __|
%   | |_| | (_) | (__| |_| | | | | | |  __/ | | | |_
%   |____/ \___/ \___|\__,_|_| |_| |_|\___|_| |_|\__|


\begin{document}

\maketitle

\pagebreak
\hspace{1pt}\vfill
\noindent
Copyright \copyright{} 2023 Carl Martin Ludvig Sinander.

\begin{quotation}
\noindent
Permission is granted to copy, distribute and/or modify this document under the terms of the \href{https://www.gnu.org/licenses/fdl}{GNU Free Documentation License}, Version 1.3 or any later version published by the Free Software Foundation; with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of the license is included in the section entitled `GNU
Free Documentation License'.
\end{quotation}

\noindent
This is a `copyleft' licence.
Visit \href{https://www.gnu.org/licenses/copyleft}{gnu.org/licenses/copyleft} to learn more.



%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Preface}
\label{preface}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

These are the notes for a second-year graduate course that I taught at Oxford in the autumns of 2021, 2022 and 2023.
Thanks to Gregorio Curello, Deniz Kattwinkel, Elliot Lipnowski, Qingmin Liu, Alex Teytelboym and to my students for comments, suggestions and insights,
and to Catharina Behrens, Josh le Cornu, Tanxin Gao, Brooklyn Han, Simon Handreke, Sanjari Kalantri, Marie Kaul, Nils Lager, Arvo Munoz Moran, Filip Tokarski, Gautam Vyas and Yidan Xu for reporting errors.



%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
% Table of contents
\pagebreak
\microtypesetup{protrusion=false}
\setcounter{tocdepth}{1}
\tableofcontents
\microtypesetup{protrusion=true}
%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%



\setcounter{chapter}{-1}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
\label{ch0}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

Much of modern economic theory concerns \emph{incentives.}
The broad question is how to get some `agents' (e.g. individuals or firms) to behave as we'd like them to.
This is achieved by designing institutions or contracts, usually abstractly called `mechanisms'.
We often imagine there being a `principal' who is tasked with designing the mechanism; but in some applications, there isn't literally a principal.

Incentive provision is sometimes trivial.
Suppose we wish that a child refrain from cursing in the classroom.
If the child's teacher is in the classroom observing her, and is able to threaten the child with the naughty corner (and the child understands and greatly fears this horrible prospect), then the child can be expected to behave.
More abstractly, incentives are easily provided when the agents can (a) be monitored accurately and (b) be rewarded generously and/or punished harshly.%
	\footnote{Such mechanisms, in which agents are kept in line using the threat of a harsh punishment, are often called `(Mirrlees) shoot-the-agent' mechanisms.}

Incentive design becomes non-trivial as soon as the principal is
%
\begin{enumerate}[label=(\alph*)]

	\item \label{bullet:info}
	imperfectly informed (i.e. monitoring is imperfect), and/or

	\item \label{bullet:stick}
	restricted in the rewards and punishments that she can dish out.

\end{enumerate}
%
The literature is conventionally divided according to \ref{bullet:info}: \emph{what} is it that the agents know and the principal doesn't?
In \emph{hidden-action} (or \emph{moral-hazard}) models, the agents take actions that the principal doesn't observe.
In \emph{hidden-information} (or \emph{adverse-selection}) models, the agents have private information from the outset (usually about their own preferences, e.g. their valuations for a good).
Mechanism design is the study of incentive provision in the latter type of model.
(There are models that feature both moral hazard and adverse selection, of course.)

Work on mechanism design is also usefully distinguished according to \ref{bullet:stick}: what \emph{incentive tools} are available to the principal?
Older work mostly focusses on incentive provision using monetary transfers,
typically restricted only by agents' ability to walk away if charged too much (`individual rationality') and/or the constraint that money cannot be \emph{taken from} agents (`limited liability').
Much of the recent literature looks at settings in which monetary transfers are highly restricted, or entirely unavailable.

A key feature of most work on mechanism design (and on incentive design more broadly) is that the principal is given the power to \emph{commit.}
In particular, she chooses a `mechanism' (or contract, or set of institutions), which specifies payoff-relevant outcomes as a function of how the agent(s) behave.
Once agent(s) have taken their actions, the principal cannot re-optimise by choosing outcomes different from those specified by the mechanism: her previous (contingent) choice binds her.
There is also a large body of work asking how incentive design changes when commitment is relaxed.

This course will sample various aspects of modern mechanism design.
I'll begin by teaching a classical topic from a modern perspective.
The rest of the course will cover topics on (or close to) the current research frontier.
Along the way, we'll encounter a number of perspectives, tools, tricks and insights; these are just as important as the details of the models and papers themselves.



%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Selling to one agent}
\label{ch1}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\input{ch1}



%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Allocating among several agents}
\label{ch2}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\input{ch2}



%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Allocating over time, without money}
\chaptermark{Allocating over time}
\label{ch3}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\input{ch3}



%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Incentivising proposal}
\label{ch4}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\input{ch4}



% %%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%
% \chapter{Hard evidence}
% \label{ch5}
% %%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%

% % \input{ch5}

% \emph{Topics:}
% revelation principle \parencite{BullWatson2007};
% randomisation and commitment have no value \parencite{GlazerRubinstein2004,GlazerRubinstein2006,Sher2011,HartKremerPerry2017,BenporathDekelLipman2019}.



%______________________________________________________________________________




%       _                               _ _
%      / \   _ __  _ __   ___ _ __   __| (_) ___ ___  ___
%     / _ \ | '_ \| '_ \ / _ \ '_ \ / _` | |/ __/ _ \/ __|
%    / ___ \| |_) | |_) |  __/ | | | (_| | | (_|  __/\__ \
%   /_/   \_\ .__/| .__/ \___|_| |_|\__,_|_|\___\___||___/
%           |_|   |_|


\begin{appendices}

\crefalias{chapter}{appsec}
\crefalias{section}{appsec}
\crefalias{subsection}{appsec}
\crefalias{subsubsection}{appsec}




%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Measure and integral}
\label{ch:meas}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

The theory of measure and (Lebesgue) integration are the foundation of modern real analysis, which is in turn the backbone of economic theory.
You do not need to know it to take this course.
But to understand what I'm saying,
it is necessary at least to know some of the basic \emph{language} of measure theory; that's what I'll cover here.

(If you'd like to pursue research in economic theory, I would advise you to learn basic measure theory.
I taught myself from \textcite{Rosenthal2006}; this book is very accessible, except that chapter 2 is harder than it needs to be, so don't get stuck there! I now prefer the first few chapters of \textcite{Folland1999}, a very beautiful book for first-year graduate students in maths. There are lots of other standard books. Efe Ok has a `measure and probability' manuscript on his website that is specifically aimed at economists, in case you find that appealing.)


Let $X$ be a non-empty set.
A \emph{$\sigma$-algebra} on $X$ is a collection of subsets of $X$ satisfying certain properties.
If $\mathcal{X}$ is a $\sigma$-algebra, we call $(X,\mathcal{X})$ a \emph{measurable space,} and call the elements of $\mathcal{X}$ the \emph{measurable subsets} of $X$.
Often the $\sigma$-algebra $\mathcal{X}$ is left partly or entirely implicit, by the way.

A \emph{measure} on a measurable space $(X,\mathcal{X})$ is a map $\mu : \mathcal{X} \to [0,\infty]$ that is countably additive:
for any countable collection $A_1,A_2,\dots$ of pairwise disjoint measurable subsets of $X$,
we have $\mu\left( \Union_{n \in \N} A_n \right) = \sum_{n \in \N} \mu(A_n)$.
The triple $(X,\mathcal{X},\mu)$ is called a \emph{measure space.}

A measurable set $A \subseteq X$ is called \emph{$\mu$-null} iff $\mu(A)=0$. If a property is holds at every $x \in X$, except possibly for $x$ belonging to a null set $A$, then that property is said to hold \emph{($\mu$-)almost everywhere,} or `($\mu$-)a.e.'.

\begin{example}
	%
	\label{example:lebesgue_measure}
	%
	The most commonly-used measure on $X=\R$ is the \emph{Lebesgue measure,}
	which is the unique measure $\lambda$ with the property that
	$\lambda([a,b]) = b-a$ for all $a<b$.
	That is, it captures the common-sense notion of \emph{length.}

	Analogously, there's Lebesgue measure on $\R^2$, which captures \emph{area,} and Lebesgue measure on $\R^3$, which captures \emph{volume,} and so on.

	The Lebesgue measure is conventionally defined on the \emph{Lebesgue $\sigma$-algebra} (whose elements are called \emph{Lebesgue sets}).
	It is often easier to work with the coarser \emph{Borel $\sigma$-algebra} (or \emph{Borel sets}); this is the smallest $\sigma$-algebra containing every open subset of $\R$.
	%
\end{example}

If $\mu$ has the further property that $\mu(X)=1$, then it is a \emph{probability measure,}
and $(X,\mathcal{X},\mu)$ is a \emph{probability space.}
In the probability context, `almost everywhere' is usually replaced with `almost sure(ly)' or `a.s.'.

Now consider a function $f : X \to Y$, where both $X$ and $Y$ are measurable spaces. For any measurable set $B \subseteq Y$, a measure on $Y$ can tell us how `large' the set $B$ is.
How large, then, is the set $A = \{ x \in X : f(x) \in B \}$ of points $x$ in $X$ that lead to a value $f(x)$ that lives in $B$?
That question only has an answer if $A$ is a measurable set!
We call a function \emph{measurable} if this question has an answer for every set $B$.
Symbolically, $f$ is measurable iff for every measurable subset $B$ of $Y$, $\{ x \in X : f(x) \in B \}$ is a measurable subset of $X$.

In modern analysis, the standard integral is the Lebesgue integral.
This is the integral used almost exclusively in economic theory, including these notes.
The integral of a measurable function $f : X \to \R$ on a measure space $(X,\mathcal{X},\mu)$ is written $\int_X f \dd \mu$.
The integral over a measurable subset $A \subseteq X$ is defined
%
\begin{equation*}
	\int_A f \dd \mu \coloneqq \int_X f \1_A \dd \mu ,
\end{equation*}
%
where $\1_A(x) \coloneqq 1$ for $x \in A$ and $\coloneqq 0$ for $x \notin A$.

On probability spaces, measurable functions are conventionally called \emph{random variables,}
and integrals are called \emph{expectations.}
One writes
%
\begin{equation*}
	\E( f ) \coloneqq \int_X f \dd \mu .
\end{equation*}

The Riemann integral has the problem that the integral of many important functions simply fails to exist; for example, any step function.
The Lebesgue integral extends the Riemann integral:
it allows many more functions (e.g. step functions) to be integrated,
while still giving the same result as the Riemann integral whenever the latter exists.

The Lebesgue integral is defined only for functions $f : X \to \R$ that are measurable.
That is a necessary condition, but it is not sufficient:
the existence of the integral requires a further condition.
(If the further condition fails, the definition of the Lebesgue integral yields the expression $\infty - \infty$, which has no meaning; therefore we agree to say that the integral does not exist in such cases.)
The integral of a function may be infinite (equal to $\infty$ or $-\infty$).
A function is called \emph{integrable} iff both (a) its integral exists, \emph{and} (b) its integral is finite.

Whereas the Riemann integral is defined only for functions $f : \R \to \R$,
the Lebesgue integral makes sense for functions $f : X \to \R$
defined on any space $X$ you like, provided it has measurable structure (i.e. is equipped with a $\sigma$-algebra).
This is very useful.




%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Convexity}
\label{ch:convexity}
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

A \emph{vector space} (or \emph{linear space}) is a set $V$ for which it makes sense to (a) add two elements of $V$ together, and (b) to multiply an element of $V$ by a scalar. (That is, in both cases, the operation yields another element of $V$.)
These `addition' and `scalar multiplication' operations have to satisfy a number of axioms that you can find on Wikipedia.

That's rather abstract, but vector spaces are all around.
The familiar example is $\R^n$, where $n \in \N$.
There is a standard definition of what it means to `add' two vectors together: it means adding up each component separately, i.e. for vectors $x,y \in \R^n$, $z = x+y$ is defined by
%
\begin{equation*}
	z_i \coloneqq x_i + y_i
	\quad \text{for each component $i \in \{1,\dots,n\}$.}
\end{equation*}
%
Similarly, scalar multiplication has a standard definition:
for a vector $x \in \R^n$ and scalar $\alpha \in \R$, $y = \alpha x$ is the vector whose $i$th entry is $y_i \coloneqq \alpha x_i$.
This makes $\R^n$ a vector space.
It is finite-dimensional (its dimension is $n$).

Much of the familiar structure of the finite-dimensional vector space $\R^n$ extends to fancier vector spaces.
Here's a useful vector space: the space of all functions $f : [0,1] \to \R$,
where addition is defined `pointwise' ($h = f+g$ is the function such that $h(x) \coloneqq f(x) + g(x)$ for every `point' $x \in [0,1]$),
and scalar multiplication is also defined `pointwise' ($g = \alpha f$ for $\alpha \in \R$ is the function such that $g(x) \coloneqq \alpha f(x)$ for every $x \in [0,1]$).
This vector space is infinite-dimensional;
indeed we can think of a function $f : [0,1] \to \R$ as a vector with uncountably many entries, one for each $x \in [0,1]$, with value $f(x)$.
(An $n$-vector $(x_1,\dots,x_n)$ may similarly be viewed as the function $f : \{1/n,2/n,\dots,(n-1)/n,1\} \to \R$ that has $f(i/n) \coloneqq x_i$ for each $i$.)

A subset $C$ of a vector space is \emph{convex} iff
for any two of its members $x,y \in C$,
every convex combination $\alpha x + (1-\alpha) y$ for $\alpha \in (0,1)$
is also a member of $C$.
In $\R^n$, you'll already be familiar with what convex sets look like.

\setcounter{example}{1}
\begin{example}
	%
	\label{example:fg}
	%
	Consider the vector space of functions $[0,1] \to \R$.
	Let $C$ be the set of all functions $h : [0,1] \to \R$ which have the form
	$f(x) = k + (2-k) x$
	for some $k \in [0,2]$.
	(Try drawing a picture to visualise $C$.)
	This is a convex set. (Prove it!)
	%
\end{example}

An element $x$ of a convex set $C$ is called an \emph{extreme point}
iff it cannot be written as a convex combination of two \emph{distinct} points in $C$. (Any point is the convex combination of itself with itself, of course!)
That is, $x \in C$ is an extreme point of $C$ iff
we cannot find $y \neq z$ in $C$ and $\alpha \in (0,1)$
such that $x = \alpha y + (1-\alpha) z$.

\setcounter{example}{1}
\begin{example}[continued]
	%
	\label{example:fg2}
	%
	Return to the previous example.
	I claim that the function $\bar{f} \equiv 2$ (constant, always equal to $2$) is an element of $C$,
	and is in fact an extreme point of $C$. (Convince yourself! A drawing will help.)
	I further claim that the function $\underline{f}$ given by $\underline{f}(x) = 2x$ is an extreme point of $C$. (Convince yourself!)
	Finally, these are the only extreme points of $C$.
	%
\end{example}

If $C$ is a convex subset of a vector space,
a function $\phi : C \to \R$ is called \emph{convex} exactly if
%
\begin{equation*}
	\phi( \alpha x + (1-\alpha) y )
	\leq \alpha \phi(x) + (1-\alpha) \phi(y)
	\quad \text{for all $x,y \in C$ and $\alpha \in (0,1)$.}
\end{equation*}
%
If the inequality is an equality (for all $x,y$ and $\alpha$)
then $\phi$ is called \emph{linear.}

It is a fact that if $C$ is a convex set and is suitably compact,
and if $\phi : C \to \R$ is a convex function and is suitably continuous,
then there is a maximiser of $\phi$ on $C$ that is an extreme point of $C$.
(Some economists call this `Bauer's maximum principle'.)
This is easy to see in $\R^2$ (visualise it!).
If $\phi$ is linear (not merely convex), then you don't have to worry about the `suitably continuous' bit, as linearity gives it to you for free.
(You \emph{do} have to worry about the `suitably compact' bit. In principle.)



\end{appendices}


%______________________________________________________________________________




%    ____  _ _     _ _                             _
%   | __ )(_) |__ | (_) ___   __ _ _ __ __ _ _ __ | |__  _   _
%   |  _ \| | '_ \| | |/ _ \ / _` | '__/ _` | '_ \| '_ \| | | |
%   | |_) | | |_) | | | (_) | (_| | | | (_| | |_) | | | | |_| |
%   |____/|_|_.__/|_|_|\___/ \__, |_|  \__,_| .__/|_| |_|\__, |
%                            |___/          |_|          |___/


% \pagebreak
\printbibliography[heading=bibintoc]



%______________________________________________________________________________



\end{document}
